# Prep AEU Data

**Purpose:** Prep AEU data for PESP merge.

**Author:** Perry ([sarah.perry\@water.ca.gov](mailto:sarah.perry@water.ca.gov){.email})

# Initial Setup
```{r, message=FALSE, warning=FALSE}
# Import Packages ---------------------------------------------------------
library(tidyverse)
library(fuzzyjoin)
library(glue)
library(lubridate)

source('admin/global_functions/global_funcs.R')
source('admin/global_functions/syn_funcs.R')
source('admin/global_functions/bsa_funcs.R')
source('admin/global_functions/check_funcs.R')
```

# Read in data
```{r}
# specify the file path
data_path_emp <- abs_pesp_path('Groups/DWR-EMP/02 Program Data/EMP_program_draft_2008-2024.csv')

# new data to format and append
df_emp <- read_quiet_csv(data_path_emp)
```

# Clean data

(Time is already in PST)

## Add in Metadata
```{r}
# Survey
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'Survey')

# Sample Method
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'SampleMethod')

# Magnification
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'Magnification')

# Depth Type
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'DepthType')

# Sample Depth
df_emp <- df_emp %>%
  mutate(SampleDepth = '1')

# Sample Scheme
df_emp <- df_emp %>%
  mutate(SampleScheme =
           case_when(str_detect(Station, 'LSZ') ~ 'random',
                     TRUE ~ 'fixed')
         )

# Count Method
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'CountMethodSample')

df_emp <- df_emp %>%
  rename(CountMethodTaxa = CountMethod) %>%
  mutate(CountMethodTaxa = stringr::str_to_lower(CountMethodTaxa))
```

## Add extra columns
```{r}
df_emp <- df_emp %>% 
  mutate(TowNetRadius = NA)
```

## Format PESP Taxon
## Extract Taxon
```{r}
df_emp <- extract_program_taxons(df_emp)
```

## Correct typos
```{r}
df_emp <- correct_taxon_typos(df_emp)
corrected_typos <- attr(df_emp, 'log')$taxon_corrections
write_log_file(corrected_typos, 'Synthesized Dataset/01 Group Checks/DWR-EMP/tables/EMP_corrected_typos.csv')

print(corrected_typos)
```

### Standardize unknowns
```{r}
df_emp <- clean_unknowns(df_emp, std_sp = TRUE, std_suffix = TRUE)
standardized_unknowns <- attr(df_emp, 'log')$clean_unknowns
write_log_file(standardized_unknowns, 'Synthesized Dataset/01 Group Checks/DWR-EMP/tables/EMP_standardized_unknowns.csv')

print(standardized_unknowns)
```

### Update synonyms
```{r}
df_emp <- update_synonyms(df_emp)
updated_synonyms <- attr(df_emp, 'log')$synonym_updates
write_log_file(updated_synonyms, 'Synthesized Dataset/01 Group Checks/DWR-EMP/tables/EMP_updated_synonyms.csv')
print(updated_synonyms)
```

### Add in higher-level taxa
```{r}
df_emp <- higher_lvl_taxa(df_emp, std_type = 'PESP')
needs_hierarchy <- attr(df_emp, 'log')$unmatched_taxa

print(needs_hierarchy)
```

## Subset relevant columns
```{r}
df_emp <- subset_cols(df_emp)
```

## Save .Rdata
```{r}
save(df_emp, file = 'synthesized/02_group_data/DWR-EMP.Rdata')
```
