# Prep AWCA Data

**Purpose:** Prep AWCA data for PESP merge.

**Author:** Perry ([sarah.perry\@water.ca.gov](mailto:sarah.perry@water.ca.gov){.email})

# Initial Setup
```{r, message=FALSE, warning=FALSE}
# Import Packages ---------------------------------------------------------
library(tidyverse)
library(fuzzyjoin)
library(glue)
library(lubridate)

source('admin/global_functions/global_funcs.R')
source('admin/global_functions/syn_funcs.R')
source('admin/global_functions/bsa_funcs.R')
source('admin/global_functions/check_funcs.R')
```

# Read in data
```{r}
# specify the file path
data_path_awca <- abs_pesp_path('Groups/DWR-AWCA/02 Program Data/AWCA_program_draft.csv')

# new data to format and append
df_awca <- read_quiet_csv(data_path_awca)
```

# Clean data

## Rename columns
```{r}
rename_map <- c(
  'Station' = 'station',
  'Datetime' = 'date_time_pdt',
  'GALD' = 'gald',
  'PhytoForm' = 'phyto_form',
  'Cells_per_mL' = 'cells_per_ml',
  'Units_per_mL' = 'organisms_per_ml',
  'Biovolume_per_mL' = 'biovolume_cubic_micron_per_ml',
  'QualityCheck' = 'quality_check',
  'Debris' = 'debris'
)

df_awca <- rename_cols(df_awca, rename_map)

df_awca <- df_awca %>%
  mutate(
    Taxon = case_when(
      !is.na(taxon_original) ~ taxon_original,
      TRUE ~ taxon_current
    )
  )
```

```{r}
unique(df_awca$SampleDepth)
```


## Extract Location
```{r}
df_awca <- df_awca %>%
  mutate(Location = str_extract(Station, '^[^_]+'))
```

## Remove treated stations
```{r}
treated <- c('DI','LH')

df_awca <- df_awca %>%
  filter(!(Location %in% treated))
```

## Format Datetime
```{r}
# Time is already PST

df_awca <- df_awca %>%
  mutate(
    Date = as.Date(Datetime),
    Time = format(Datetime, '%H:%M:%S')
  )
```

## Add Lat/Lon
```{r}
df_awca <- add_latlon(df_awca, 'Groups/DWR-AWCA/Metadata/Stations-AWCA.csv')
```

## Add in Metadata
```{r}
# Survey
df_awca <- add_meta_col(df_awca, 'DWR-AWCA', 'Survey')

# Magnification
df_awca <- add_meta_col(df_awca, 'DWR-AWCA', 'Magnification')

# Sample Scheme
df_awca <- add_meta_col(df_awca, 'DWR-AWCA', 'SampleScheme')

# Sample Method
df_awca <- add_meta_col(df_awca, 'DWR-AWCA', 'SampleMethod')

# Depth Type
df_awca <- add_meta_col(df_awca, 'DWR-AWCA', 'DepthType')

# Sample Depth
df_awca <- add_meta_col(df_awca, 'DWR-AWCA', 'SampleDepth')

#Lab
df_awca <- add_meta_col(df_awca, 'DWR-AWCA', 'Lab')

# Notes
df_awca <- df_awca %>% mutate(Notes = 'Unknown')

# Count Method
df_awca <- add_meta_col(df_awca, 'DWR-AWCA', 'CountMethodSample')

df_awca <- df_awca %>%
  mutate(CountMethodTaxa = CountMethodSample) # must be the same

# convert "NA" to NA
df_awca <- df_awca %>%
  mutate(across(where(is.character), ~na_if(.x, 'NA')))
```

## Add extra columns
```{r}
df_awca <- df_awca %>% 
  mutate(TowNetRadius = NA,
         OrigTaxon = Taxon)
```

## Remove old taxa info
```{r}
df_awca <- remove_taxa_info(df_awca)
```

## Add PESP Taxon
## Correct typos
```{r}
df_awca <- correct_taxon_typos(df_awca)
corrected_typos <- attr(df_awca, 'log')$taxon_corrections
write_log_file(corrected_typos, 'Synthesized Dataset/01 Group Checks/DWR-AWCA/tables/AWCA_corrected_typos.csv')

print(corrected_typos)
```

### Standardize unknowns
```{r}
df_awca <- clean_unknowns(df_awca, std_sp = TRUE, std_suffix = TRUE)
standardized_unknowns <- attr(df_awca, 'log')$clean_unknowns
write_log_file(standardized_unknowns, 'Synthesized Dataset/01 Group Checks/DWR-AWCA/tables/AWCA_standardized_unknowns.csv')

print(standardized_unknowns)
```

### Update synonyms
```{r}
df_awca <- update_synonyms(df_awca)
updated_synonyms <- attr(df_awca, 'log')$synonym_updates
write_log_file(updated_synonyms, 'Synthesized Dataset/01 Group Checks/DWR-AWCA/tables/AWCA_updated_synonyms.csv')

print(updated_synonyms)
```

### Add in higher-level taxa
```{r}
df_awca <- higher_lvl_taxa(df_awca, std_type = 'pesp')
needs_hierarchy <- attr(df_awca, 'log')$unmatched_taxa

print(needs_hierarchy)
```

## Add ID column
```{r}
df_awca <- add_id_col(df_awca, loc_col = 'Station')
```

## Subset relevant columns
```{r}
df_awca <- subset_cols(df_awca)

df_awca <- df_awca %>%
  arrange(Date, Time) %>%
  mutate(
    Cells_per_mL = round(Cells_per_mL, 2),
    Units_per_mL = round(Units_per_mL, 2),
    Biovolume_per_mL = round(Biovolume_per_mL, 2)
  )
```

# Rename QC col codes
```{r}
df_awca <- df_awca %>%
  mutate(Notes =
           case_when(QualityCheck == 'fragmented' ~ 'FragmentedDiatoms',
                     TRUE ~ 'Unknown')
         ) %>%
  mutate(QualityCheck =
           case_when(QualityCheck == 'good' ~ 'NoCode',
                     QualityCheck == 'fragmented' ~ 'NoCode',
                     QualityCheck == 'degraded' ~ 'Degraded',
                     TRUE ~ QualityCheck)
         ) %>%
  mutate(Debris =
           case_when(Debris == 'high' ~ 'High',
                     Debris == 'moderate' ~ 'Moderate',
                     Debris == 'low' ~ 'Low',
                     is.na(Debris) ~ 'None',
                     TRUE ~ Debris))

unique(df_awca$Debris)
unique(df_awca$Notes)
unique(df_awca$QualityCheck)
```

## Combine taxa (multiple sizes, standardized unknowns)
```{r}
df_awca <- combine_taxons(df_awca, c('Date','Station','Time'))
combine_issues <- attr(df_awca, 'log')$combined_conflicts
combine_checks <- attr(df_awca, 'log')$combined_taxa

combine_checks
```

# Save .Rdata
```{r}
save(df_awca, file = 'synthesized/02_group_data/DWR-AWCA.Rdata')
```
