# Prep AEU Data

**Purpose:** Prep AEU data for PESP merge.

**Author:** Perry ([sarah.perry\@water.ca.gov](mailto:sarah.perry@water.ca.gov){.email})

# Initial Setup
```{r, message=FALSE, warning=FALSE}
# Import Packages ---------------------------------------------------------
library(tidyverse)
library(fuzzyjoin)
library(glue)
library(lubridate)

source('admin/global_functions/global_funcs.R')
source('programs/USBR-DOP/ref_code/DOP_funcs.R')
source('admin/global_functions/syn_funcs.R')
source('admin/global_functions/bsa_funcs.R')
source('admin/global_functions/check_funcs.R')
```

# Read in data
```{r}
# specify the file path
data_path_dop <- abs_pesp_path('Groups/USBR-DOP/02 Program Data/DOP_program_draft.csv')

# new data to format and append
df_dop <- read_quiet_csv(data_path_dop)
```

# Clean data

## Convert Time
```{r}
df_dop <- convert_to_pst(df_dop)

converted_times <- attr(df_dop, 'log')$converted_times
write_log_file(converted_times, 'Synthesized Dataset/01 Group Checks/USBR-DOP/tables/DOP_converted_times.csv')
```

## Add in Metadata
```{r}
# Survey
df_dop <- add_meta_col(df_dop, 'USBR-DOP', 'Survey')

# Magnification
df_dop <- add_meta_col(df_dop, 'USBR-DOP', 'Magnification')

# SampleScheme
df_dop <- add_meta_col(df_dop, 'USBR-DOP', 'SampleScheme')

# Depth Type
df_dop <- add_meta_col(df_dop, 'USBR-DOP', 'DepthType')

# Sample Depth
df_dop <- add_meta_col(df_dop, 'USBR-DOP', 'SampleDepth')

# Count Method
df_dop <- add_meta_col(df_dop, 'USBR-DOP', 'CountMethodSample')

df_dop <- df_dop %>%
  mutate(CountMethodTaxa = CountMethodSample)

# make 'Station'
df_dop <- df_dop %>%
  rename(Station = Sortable_ID) %>%
  mutate(Location = Station)
```

## Add extra columns
```{r}
df_dop <- df_dop %>% 
  mutate(TowNetRadius = NA,
         Units_per_mL = NA,
         Biovolume_per_mL = NA,
         GALD = NA,
         PhytoForm = NA,
         QualityCheck = 'Unknown',
         Debris = 'Unknown',
         Notes = 'Unknown')
```

## Format PESP Taxon
## Extract Taxon
```{r}
df_dop <- extract_program_taxons(df_dop)
```

## Correct typos
```{r}
df_dop <- correct_taxon_typos(df_dop)
corrected_typos <- attr(df_dop, 'log')$taxon_corrections
write_log_file(corrected_typos, 'Synthesized Dataset/01 Group Checks/USBR-DOP/tables/DOP_corrected_typos.csv')

print(corrected_typos)
```

### Standardize unknowns
```{r}
df_dop <- clean_unknowns(df_dop, std_sp = TRUE, std_suffix = TRUE)
standardized_unknowns <- attr(df_dop, 'log')$clean_unknowns
write_log_file(standardized_unknowns, 'Synthesized Dataset/01 Group Checks/USBR-DOP/tables/DOP_standardized_unknowns.csv')

print(standardized_unknowns)
```

### Update synonyms
```{r}
df_dop <- update_synonyms(df_dop)
updated_synonyms <- attr(df_dop, 'log')$synonym_updates
write_log_file(updated_synonyms, 'Synthesized Dataset/01 Group Checks/USBR-DOP/tables/DOP_updated_synonyms.csv')
print(updated_synonyms)
```

### Add in higher-level taxa
```{r}
df_dop <- higher_lvl_taxa(df_dop, std_type = 'PESP')
needs_hierarchy <- attr(df_dop, 'log')$unmatched_taxa

print(needs_hierarchy)
```

## Add ID column
```{r}
df_dop <- add_id_col(df_dop)
```

## Subset relevant columns
```{r}
df_dop <- subset_cols(df_dop)
```

## Combine taxa (multiple sizes, standardized unknowns)
```{r}
df_dop <- combine_taxons(df_dop)
combine_checks <- attr(df_dop, 'log')$combined_conflicts
```

## Save .Rdata
```{r}
save(df_dop, file = 'synthesized/02_group_data/USBR-DOP.Rdata')
```
