# Format original AEU data

**Purpose:** Code to format AEU data for original PESP upload.

**Author:** Perry ([sarah.perry\@water.ca.gov](mailto:sarah.perry@water.ca.gov){.email})

# Initial Setup
```{r, message=FALSE, warning=FALSE}
# Import Packages ---------------------------------------------------------
library(tidyverse)
library(fuzzyjoin)
library(glue)
library(lubridate)

source('admin/global_functions/global_funcs.R')
source('programs/DWR-AEU/code/aeu_funcs.R')
source('admin/global_functions/bsa_funcs.R')
source('admin/global_functions/check_funcs.R')
```

## Read in data
```{r}
# specify the file path
data_path <- abs_pesp_path('Groups/DWR-AEU/01 Raw Data/dwr-aeu-2020.csv')

# new data to format and append
df_data <- read_quiet_csv(data_path)
```

## Check distinct columns
```{r}
df_data <- check_distinct(df_data, type = 'full')
nondistinct_rows <- attr(df_data, 'log')$nondistinct_allrows
write_log_file(nondistinct_rows, 'Groups/DWR-AEU/02 Processed Data/QC files/tables/AEU_nondistinct_rows.csv')

# remove duplicates
df_data <- df_data %>% distinct()

print(nondistinct_rows)
```


## Standardize data and add metadata columns
```{r}
df_data <- standardize_cols_aeu(df_data)

df_data <- add_meta_col(df_data, 'Yolo Bypass Fish Monitoring Program', 'SampleDepth')

df_data <- add_meta_col(df_data, 'Yolo Bypass Fish Monitoring Program', 'Lab')

df_data <- add_meta_col(df_data, 'Yolo Bypass Fish Monitoring Program', 'SampleMethod')
```

# Add Taxonomy (and QC column)

## Correct typos
```{r}
df_data <- correct_taxon_typos(df_data)
corrected_typos <- attr(df_data, 'log')$taxon_corrections
write_log_file(corrected_typos, 'Groups/DWR-AEU/02 Processed Data/QC files/tables/AEU_corrected_typos.csv')

print(corrected_typos)
```

## Add QC column
```{r}
df_data <- add_qc_col(df_data)
df_data <- add_debris_col(df_data)

df_data <- extract_unstandardized_comments(df_data, 'Comments', delimiter = '. ')
unmatched_comments <- attr(df_data, 'log')$unmatched_comments
write_log_file(unmatched_comments, 'Groups/DWR-AEU/02 Processed Data/QC files/tables/AEU_leftover_comments.csv')

print(unmatched_comments)
```

## Standardize unknowns
```{r}
df_data <- clean_unknowns(df_data)
standardized_unknowns <- attr(df_data, 'log')$clean_unknowns
write_log_file(standardized_unknowns, 'Groups/DWR-AEU/02 Processed Data/QC files/tables/AEU_standardized_unknowns.csv')

print(standardized_unknowns)
```

## Update synonyms
```{r}
df_data <- update_synonyms(df_data)
updated_synonyms <- attr(df_data, 'log')$synonym_updates
write_log_file(updated_synonyms, 'Groups/DWR-AEU/02 Processed Data/QC files/tables/AEU_updated_synonyms.csv')

print(updated_synonyms)
```

## Add in higher-level taxa
```{r}
df_data <- higher_lvl_taxa(df_data)
needs_hierarchy <- attr(df_data, 'log')$unmatched_taxa
write_log_file(needs_hierarchy, 'Groups/DWR-AEU/02 Processed Data/QC files/tables/AEU_needs_hierarchy.csv')

print(needs_hierarchy)
```

## Check distinct cols before combining
```{r}
df_data <- check_distinct(df_data, type = 'full')
attr(df_data, 'log')$nondistinct_allrows

df_data <- check_distinct(df_data, type = 'key_cols')
attr(df_data, 'log')$nondistinct_keyrows
```

## Combine taxa (multiple sizes, standardized unknowns)
```{r}
df_data <- combine_taxons(df_data)
combined_taxa <- attr(df_data, 'log')$combined_taxa
write_log_file(combined_taxa, 'Groups/DWR-AEU/02 Processed Data/QC files/tables/AEU_combined_taxa.csv')

print(combined_taxa)
```

# Finish formatting

## Subset columns
```{r}
# subset relevant columns
df_data <- subset_cols_aeu(df_data)
```

## Add different dilution factor
```{r}
df_data <- add_dilution_qc(df_data)
```

## Checks
```{r}
# check for NAs
df_data <- check_nas(df_data)
na_check <- attr(df_data, 'log')$na_check

print(na_check)
```

# Export
```{r}
write_csv(df_data, abs_pesp_path('Groups/DWR-AEU/02 Processed Data/AEU_draft.csv'))
```
