# Format original AEU data

**Purpose:** Code to format AEU data for original PESP upload.

**Author:** Perry ([sarah.perry\@water.ca.gov](mailto:sarah.perry@water.ca.gov){.email})

# Initial Setup

```{r, message=FALSE, warning=FALSE}
# Import Packages ---------------------------------------------------------
library(tidyverse)
library(fuzzyjoin)
library(glue)
library(lubridate)

source('admin/global_functions/global_funcs.R')
source('programs/DWR-AEU/code/aeu_funcs.R')
source('admin/global_functions/bsa_funcs.R')
source('admin/global_functions/check_funcs.R')
```

## Read in data
```{r}
# specify the file path
data_path <- abs_pesp_path('Groups/DWR-AEU/01 Raw Data/dwr-aeu-2020.csv')

# new data to format and append
df_data <- read_quiet_csv(data_path)
```

# Check distinct columns
```{r}
df_data <- check_distinct(df_data, type = 'full')
attr(df_data, 'log')$nondistinct_allrows

# remove duplicates
df_data <- df_data %>% distinct()
```


## Standardize data and add metadata columns
```{r}
df_data <- standardize_cols_aeu(df_data)

df_data <- add_meta_col(df_data, 'Yolo Bypass Fish Monitoring Program', 'SamplingDepth')

df_data <- add_meta_col(df_data, 'Yolo Bypass Fish Monitoring Program', 'Lab')

df_data <- add_meta_col(df_data, 'Yolo Bypass Fish Monitoring Program', 'SamplingMethod')
```

# Add Taxonomy (and QC column)

## Correct typos
```{r}
df_data <- correct_taxon_typos(df_data)
attr(df_data, 'log')$taxon_corrections
```

## Add QC column
```{r}
df_data <- add_qc_col(df_data)
df_data <- add_debris_col(df_data)

df_data <- extract_unstandardized_comments(df_data, 'Comments', delimiter = '. ')
attr(df_data, 'log')$unmatched_comments
```

## Standardize unknowns
```{r}
df_data <- clean_unknowns(df_data)
attr(df_data, 'log')$clean_unknowns
```

## Update synonyms
```{r}
df_data <- update_synonyms(df_data)
attr(df_data, 'log')$synonym_updates
```

## Add in higher-level taxa
```{r}
df_data <- higher_lvl_taxa(df_data)
attr(df_data, 'log')$unmatched_taxa
```

## Check distinct cols before combining
```{r}
df_data <- check_distinct(df_data, type = 'full')
attr(df_data, 'log')$nondistinct_allrows

df_data <- check_distinct(df_data, type = 'key_cols')
attr(df_data, 'log')$nondistinct_keyrows
```

## Combine taxa (multiple sizes, standardized unknowns)
```{r}
df_data <- combine_taxons(df_data)
attr(df_data, 'log')$combined_taxa
```

# Finish formatting
```{r}
df_data <- subset_cols_aeu(df_data)
```

```{r}
# add different dilution code for data before/including 2020
df_data <- add_dilution_qc(df_data)
```

```{r}
# # Check Collection Type Variables -----------------------------------------------
check_methods(df_data)
```
## Export

```{r}
# write_csv(df_joined, 'programs/DWR-AEU/data/phyto_data_AEU.csv')
```
