# Prep FRP Data

**Purpose:** Prep FRP data for PESP merge.

**Author:** Perry ([sarah.perry\@water.ca.gov](mailto:sarah.perry@water.ca.gov){.email})

# Initial Setup
```{r, message=FALSE, warning=FALSE}
# Import Packages ---------------------------------------------------------
library(tidyverse)
library(fuzzyjoin)
library(glue)
library(lubridate)

source('admin/global_functions/global_funcs.R')
source('admin/global_functions/syn_funcs.R')
source('admin/global_functions/bsa_funcs.R')
source('admin/global_functions/check_funcs.R')
```

# Read in data
```{r}
# specify the file path
data_path_frp <- abs_pesp_path('Groups/CDFW-FRP/02 Processed Data/FRP_PESP_updated.csv')

# new data to format and append
df_frp <- read_quiet_csv(data_path_frp)
```

# Clean data

## Rename columns
```{r}
rename_map <- c(
  'Station' = 'VisitNo',
  'Latitude' = 'LatitudeStart',
  'Longitude' = 'LongitudeStart',
  'Time' = 'StartTime',
  'SampleDepth' = 'DepthOfSample',
  'Lab' = 'LAB_NAME',
  'Taxon'= 'OriginalTaxon',
  'Cells_per_mL' = 'Cells_permL',
  'Units_per_mL' = 'Units_permL',
  'Biovolume_per_mL' = 'Biov_permL'
)

df_frp <- rename_cols(df_frp, rename_map)
```

## Convert Time (if needed)
TODO: come back to this when metadata is in

## Add DepthType

```{r}
df_frp <- df_frp %>%
  mutate(DepthType =
           case_when(Substrate == 'epiphytic' ~ 'surface',
                     Substrate == 'pelagic' ~ 'near surface',
                     Substrate == 'benthic' ~ 'bottom'),
         SampleMethod =
           case_when(Substrate == 'epiphytic' ~ 'putty knife',
                     Substrate == 'pelagic' & Date < '2023-07-01' ~ 'unspecified surface grab',
                     Substrate == 'pelagic' & Date >= '2023-07-01' ~ 'Van Dorn',
                     Substrate == 'benthic' ~ 'syringe'),
         SampleDepth =
           case_when(SampleMethod == 'Van Dorn' ~ 0.5,
                     SampleDepth == 0 | SampleDepth == 0.0 ~ NA,
                     TRUE ~ SampleDepth)
         )
```


## Add in Metadata
```{r}
# Survey
df_frp <- add_meta_col(df_frp, 'CDFW-FRP', 'Survey')

# Magnification
df_frp <- add_meta_col(df_frp, 'CDFW-FRP', 'Magnification')

# Sample Scheme
df_frp <- add_meta_col(df_frp, 'CDFW-FRP', 'SampleScheme')
```

## Add extra columns
```{r}
df_frp <- df_frp %>% 
  mutate(TowNetRadius = NA,
         OrigTaxon = Taxon)
```


## Add PESP Taxon
## Standardize unknowns

```{r}
df_frp <- clean_unknowns(df_frp)
standardized_unknowns <- attr(df_frp, 'log')$clean_unknowns

print(standardized_unknowns)
```

## Update synonyms

```{r}
df_frp <- update_synonyms(df_frp)
updated_synonyms <- attr(df_frp, 'log')$synonym_updates
print(updated_synonyms)
```

## Add in higher-level taxa

```{r}
df_frp <- higher_lvl_taxa(df_frp)
needs_hierarchy <- attr(df_frp, 'log')$unmatched_taxa

print(needs_hierarchy)
```


## Subset relevant columns
```{r}
df_frp <- subset_cols(df_frp)
```

# Save .Rdata
```{r}
save(df_frp, file = 'synthesized/group_prep/group_data/CDFW-FRP.Rdata')
```

