# Prep AEU Data

**Purpose:** Prep AEU data for PESP merge.

**Author:** Perry ([sarah.perry\@water.ca.gov](mailto:sarah.perry@water.ca.gov){.email})

# Initial Setup
```{r, message=FALSE, warning=FALSE}
# import packages
devtools::load_all()
```

# Read in data
```{r}
# specify the file path
data_path_aeu <- abs_pesp_path('Groups/DWR-AEU/02 Program Data/AEU_program_draft.csv')

# new data to format and append
df_aeu <- read_quiet_csv(data_path_aeu)
```

# Clean data

## Convert Time
```{r}
df_aeu <- convert_to_pst(df_aeu)

converted_times <- attr(df_aeu, 'log')$converted_times
write_log_file(converted_times, 'Synthesized Dataset/01 Group Checks/DWR-AEU/tables/AEU_converted_times.csv')
```

## Add in Metadata
```{r}
# Survey
df_aeu <- add_meta_col(df_aeu, 'DWR-YBFMP', 'Survey')

# Sample Method
df_aeu <- add_meta_col(df_aeu, 'DWR-YBFMP', 'SampleMethod')

# Magnification
df_aeu <- add_meta_col(df_aeu, 'DWR-YBFMP', 'Magnification')

# Depth Type
df_aeu <- add_meta_col(df_aeu, 'DWR-YBFMP', 'DepthType')

# Sample Scheme
df_aeu <- add_meta_col(df_aeu, 'DWR-YBFMP', 'SampleScheme')

# Count Method
df_aeu <- add_meta_col(df_aeu, 'DWR-YBFMP', 'CountMethodSample')

df_aeu <- df_aeu %>%
  mutate(CountMethodTaxa = CountMethodSample) # must be the same
```

## Add extra columns
```{r}
df_aeu <- df_aeu %>% 
  mutate(TowNetRadius = NA,
         Location = Station)
```

## Format PESP Taxon
## Extract Taxon
```{r}
df_aeu <- extract_program_taxa(df_aeu)
```

## Correct typos
```{r}
df_aeu <- correct_taxon_typos(df_aeu)
corrected_typos <- attr(df_aeu, 'log')$taxon_corrections
write_log_file(corrected_typos, 'Synthesized Dataset/01 Group Checks/DWR-AEU/tables/AEU_corrected_typos.csv')

print(corrected_typos)
```

### Standardize unknowns
```{r}
df_aeu <- clean_unknowns(df_aeu, std_sp = TRUE, std_suffix = TRUE)
standardized_unknowns <- attr(df_aeu, 'log')$clean_unknowns
write_log_file(standardized_unknowns, 'Synthesized Dataset/01 Group Checks/DWR-AEU/tables/AEU_standardized_unknowns.csv')

print(standardized_unknowns)
```

### Update synonyms
```{r}
df_aeu <- update_synonyms(df_aeu)
updated_synonyms <- attr(df_aeu, 'log')$synonym_updates
write_log_file(updated_synonyms, 'Synthesized Dataset/01 Group Checks/DWR-AEU/tables/AEU_updated_synonyms.csv')
print(updated_synonyms)
```

### Add in higher-level taxa
```{r}
df_aeu <- higher_lvl_taxa(df_aeu, std_type = 'PESP')
needs_hierarchy <- attr(df_aeu, 'log')$unmatched_taxa

print(needs_hierarchy)
```

## Add ID column
```{r}
df_aeu <- add_id_col(df_aeu, loc_col = 'Station')
```

## Subset relevant columns
```{r}
df_aeu <- subset_cols(df_aeu)
```

## Combine taxa (multiple sizes, standardized unknowns)
```{r}
df_aeu <- combine_taxa(df_aeu)
combine_checks <- attr(df_aeu, 'log')$combined_conflicts
```

## Save .Rdata
```{r}
save(df_aeu, file = 'admin/standardized_data/DWR-AEU.Rda')
```
