# Prep AEU Data

**Purpose:** Prep AEU data for PESP merge.

**Author:** Perry ([sarah.perry\@water.ca.gov](mailto:sarah.perry@water.ca.gov){.email})

# Initial Setup
```{r, message=FALSE, warning=FALSE}
# import packages
devtools::load_all()
```

# Read in data
```{r}
# specify the file path
data_path_emp <- abs_pesp_path('Groups/DWR-EMP/02 Program Data/EMP_program_draft_2008-2024.csv')

# new data to format and append
df_emp <- read_quiet_csv(data_path_emp)
```

# Clean data

(Time is already in PST)

## Add in Metadata
```{r}
# Survey
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'Survey')

# Sample Method
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'SampleMethod')

# Magnification
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'Magnification')

# Depth Type
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'DepthType')

# Sample Depth
df_emp <- df_emp %>%
  mutate(SampleDepth = '1')

# Sample Scheme
df_emp <- df_emp %>%
  mutate(SampleScheme =
           case_when(str_detect(Station, 'LSZ') ~ 'random',
                     TRUE ~ 'fixed')
         )

# Count Method
df_emp <- add_meta_col(df_emp, 'DWR-EMP', 'CountMethodSample')

df_emp <- df_emp %>%
  rename(CountMethodTaxa = CountMethod) %>%
  mutate(CountMethodTaxa = stringr::str_to_lower(CountMethodTaxa))
```

## Add extra columns
```{r}
df_emp <- df_emp %>% 
  mutate(TowNetRadius = NA)
```

## Remove TB, set LSZ locations, and remove BadData
```{r}
df_emp <- df_emp %>%
  filter(Station != 'TB')

df_emp <- df_emp %>%
  mutate(
    Location = case_when(
      Station == 'LSZ2' ~ 'LSZ2',
      Station == 'LSZ6' ~ 'LSZ6',
      Station == 'LSZ2-SJR' ~ 'LSZ2-SJR',
      Station == 'LSZ6-SJR' ~ 'LSZ6-SJR',
      TRUE ~ Location
    )
  )

df_emp <- df_emp %>%
  filter(QualityCheck != 'BadData')
```


## Format PESP Taxon
## Extract Taxon
```{r}
df_emp <- extract_program_taxa(df_emp)
```

## Correct typos
```{r}
df_emp <- correct_taxon_typos(df_emp)
corrected_typos <- attr(df_emp, 'log')$taxon_corrections
write_log_file(corrected_typos, 'Synthesized Dataset/01 Group Checks/DWR-EMP/tables/EMP_corrected_typos.csv')

print(corrected_typos)
```

### Standardize unknowns
```{r}
df_emp <- clean_unknowns(df_emp, std_sp = TRUE, std_suffix = TRUE)
standardized_unknowns <- attr(df_emp, 'log')$clean_unknowns
write_log_file(standardized_unknowns, 'Synthesized Dataset/01 Group Checks/DWR-EMP/tables/EMP_standardized_unknowns.csv')

print(standardized_unknowns)
```

### Update synonyms
```{r}
df_emp <- update_synonyms(df_emp)
updated_synonyms <- attr(df_emp, 'log')$synonym_updates
write_log_file(updated_synonyms, 'Synthesized Dataset/01 Group Checks/DWR-EMP/tables/EMP_updated_synonyms.csv')
print(updated_synonyms)
```

### Add in higher-level taxa
```{r}
df_emp <- higher_lvl_taxa(df_emp, std_type = 'PESP')
needs_hierarchy <- attr(df_emp, 'log')$unmatched_taxa

print(needs_hierarchy)
```

## Add ID column
```{r}
df_emp <- add_id_col(df_emp, loc_col = 'Station')
```

## Subset relevant columns
```{r}
df_emp <- subset_cols(df_emp)
```

## Combine taxa (multiple sizes, standardized unknowns)
```{r}
df_emp <- combine_taxa(df_emp, key_cols = c('Date','Station','CountMethodSample','CountMethodTaxa'))
combine_checks <- attr(df_emp, 'log')$combined_taxa
combine_checks
```
```{r}
df_dupes <- df_emp %>%
  group_by(Event_ID, Kingdom, Phylum, Class, AlgalGroup, Genus, Species, Taxon,CountMethodTaxa) %>%
  filter(n()  > 1) %>%
  ungroup()

df_dupes
```

## Remove rows with no data
```{r}
df_emp <- remove_nodata(df_emp)
no_data <- attr(df_emp, 'log')$nodata
```

## Check for outliers
```{r}
# Cells per mL
df_emp <- flag_outliers(df_emp, Cells_per_mL)
```

```{r}
# Units per mL
df_emp <- flag_outliers(df_emp, Units_per_mL)
```

```{r}
# Biovolume per mL
df_emp <- flag_outliers(df_emp, Biovolume_per_mL)
```

## Save .Rdata
```{r}
save(df_emp, file = 'admin/standardized_data/DWR-EMP.Rda')
```
